\section{Sorting algorithms}

Given an array $A$ with pairwise distinct elements we want to compute an array $B$ with the same elements in sorted order. There is a lower bound of $\Omega(\log n)$ on CREW and $\Omega(\sqrt{\log n})$ on CRCW machines.

\subsection{Quicksort}

Everybody knows sequential quicksort. It can be trivially parallelized by doing the recursive calls in parallel. It is more tricky to split the array in parallel.

Obviously we can compare each element to the pivot in parallel. If we construct an array that contains a 1 at position $i$ if the $i$-th element is larger than the pivot, we can compute a prefix sum over that and get the positions of the elements in the new array. With that information we can copy the values to the new array in parallel too.

\begin{lstlisting}
SPLIT
for $i=1\ldots n$ do in parallel
	if A[i] $\leq$ pivot then
		F[i] = 0
	else
		F[i] = 1

P := SCAN(F)

for $i=1\ldots n$ do in parallel
	if A[i] $\leq$ pivot then
		A$_0$[i-P[i]] = A[i]
	else
	  A$_1$[P[i]] = A[i]
\end{lstlisting}

This way the work for sorting is still $O(n\log n)$ (with a slightly larger factor), but the depth of the algorithm is only O($\log^2 n$) (in the average case).

\subsection{Mergesort}

Mergesort can also be trivially parallelized in the recursive steps, but merging in parallel is more tricky.

Suppose we have two sorted arrays $X,Y$ and want to merge them into one sorted array. The idea is to subdivide $X$ and $Y$ into a bunch of subarrays of equal size and recombine them later

\begin{lstlisting}
MERGE(X,Y)
k:= [see later]
divide X,Y into k subarrays of equal size
let the endpoints be $x_0,\ldots, x_k, y_0,\ldots , y_k$

for $i=0\ldots k$ do in parallel
	$x_i^*$ = position of $x_i$ in Y
	$y_i^*$ = position of $y_i$ in X
	
//here we have 2k subarrays for X and Y

for $i=1\ldots 2k$ do in parallel
	Z$_i$ = SequentialMerge(X$_i$, Y$_i$)
	
return Z
\end{lstlisting}

The work is 

\[W=2(k+1)O(\log n) +  O(\underbrace{\sum_{i=1}^{2k} |X_i| + |Y_i|}_{2n})\]

and the depth 

\[D=O(\log n) +O(\underbrace{\max_{i=1\ldots 2k} (X_i+Y_i)}_{\leq n/k})\]

We need to think about a good value for $k$ to get a good runtime. Obviously $k= n/\log n$ is a good choice.

\section{Pointer jumping}

Assume we have a rooted tree given as a pointer array (each entry points to its parent, the root points to itself). We want to compute a pointer from each node to the root.

\begin{lstlisting}
POINT-TO-ROOT
for $k= 1\ldots \log n$ 
	for each vertex do in parallel
		temp[v] = P[P[v]]
		P[v] = temp[v]
\end{lstlisting}

\begin{thm} Point to root is correct and has a depth of $\log n$\end{thm}
