\subsection{Colouring Trees}

In order to find a faster algorithm we first look at a special case, trees. A tree is an acyclic connected graph. All trees are bipartite, and hence have chromatic number 2.

\begin{thm} $T$ is a tree implies $\chi(T)=2$ \end{thm}

\begin{pr} Let $v$ be a distinguished vertex, \emph{the root}, in the tree. Colour $v$ red.  Colour every other vertex $w$ according to the distance from $v$. If the distance is odd colour it blue, else red. Since there are no cycles in the graph, the coloring is valid (i.e. it can't happen that a node is on two paths to the root, one of even and one of odd length).
\end{pr}

If we assume that the root knows that it's special we can construct the algorithm in figure \ref{alg:slow_tree_colour}

\begin{figure}[hbt]
\begin{lstlisting}
root: colour with red; send to all children
all except root: 
	wait for message
	if parent red
		colour blue
	else
		color red
	send colour to all children
\end{lstlisting}
\caption{Slow Tree Coloring}
\label{alg:slow_tree_colour}
\end{figure}


This algorithm isn't faster than the previous one, for degenerate trees (path). We also need to find a way to decide which node is the root. The latter problem is deferred to a later lecture.

Note that this algorithm doesn't need to operate in a synchronous environment. Every node waits for an event before it starts its computation. This is an example of an event-driven algorithm that can be implemented in an asynchronous setting.

\begin{Def}[Asynchronous DA] An asynchronous distributed algorithm has no access to a global clock and works event-driven.

Any sent message will reach its destination in finite time.
\end{Def}

There are different ways to assess the complexity of asynchronous algorithms (there are no bounds on the time a message takes). Typically we are interested in 

\begin{itemize}
\item Message Complexity: The number of messages sent
\item Time Complexity: Under the assumption that each message has delay 1, the time complexity is the length of the longest chain of message dependencies. Alternatively: the time needed if every message arrives after \emph{at most} one unit of time.
\end{itemize}

\begin{thm} The asynchronous time complexity of algorithm \ref{alg:slow_tree_colour} is $\leq \text{height}(T)$ \end{thm}
\begin{pr} In the first round the root sends its colour to all neighbours. This costs one time unit. Argument follows from induction on the height\end{pr}

We now try to improve the algorithm to get a sublinear running time. In fact we'll give an algorithm that is bounded by $O(\log^*n)$, where $\log^* n = \min \{i|\log^{(i)} n \leq 2\}$. We assume that all vertices know that the graph is a tree, a root is known and each node knows its parent. How we find that information is again deferred to later

\begin{figure}[hbt]
\begin{lstlisting}
Let $c_v$ be ID(v)
repeat 
	send $c_v$ to all children
	if $v$ is the root
		I=0
	else 
		let $c_p$ be the parent's colour
		I = smallest i s.t. bin$_i$($c_v$), bin$_i$($c_p$) differ
		
	$c_v$ = <bin(I),bin$_I$($c_v$)> //use $1+\log |c_v|$ bits  
											  //even if I small
until |bin($c_v$)| does not change
\end{lstlisting}
\caption{6-colour}
\label{alg:6-colour}
\end{figure}

\begin{lem} Algorithm \ref{alg:6-colour} produces a valid colouring \end{lem}

\begin{pr} Since all IDs are different the colouring is valid in the first step. Now consider some iteration during the execution of the algorithm

Let $u$ be the parent of $w$. If $u$ finds $I_u$ and $w$ finds $I_w$ then the new colours are $c_u=<bin(I_u), bin_{I_u}(c_u)>, c_w=<bin(I_w),b_{I_w}(c_w)>$. There are two cases

\begin{itemize}
\item $I_u\neq I_w$. Then the first part of the new colours differs. \ok
\item $I_u=I_w=I$. Then we have $b_I(c_u)\neq b_I(c_w)$, by the choice of $I$ \ok
\end{itemize}

So the colouring stays valid after each round.
\end{pr}

\begin{lem} Let $K_i$ be the number of bits in the representations of the colours in round $i$.

\begin{align*}
K_0 &= \left\lceil \log n\right\rceil\\
K_{i+1} &= 1+\left\lceil\log K_i\right\rceil
\end{align*}

\end{lem}

\begin{pr}
We have that $K_{i+1} < K_i$ as long as $K_i\geq 4$. If $K_i\in \{1,2,3\}$, we have $K_{i+1}=K_i$ (by direct calculation \ok). Otherwise we can write $K_i=2^x\cdot y$, $x\in \N$, $y\in [1,2)$. Then we have

\[K_{i+1} = 1+\left\lceil \log(2^x\cdot y)\right\rceil = \begin{cases} 1+x & y=1\\ 2+x &y>1\end{cases}\]

It follows that

\[K_{i+1}<K_i \Leftrightarrow 2^x\cdot y > \begin{cases} 1+x & y=1\\ 2+x &y>1\end{cases}\]

Check yourself that the exponential function on the left grows sufficiently fast for the claim to be true.
\end{pr}

\begin{thm} The final colouring uses at most 6 colours and stops after $O(\log^* n)$ rounds.\end{thm}

\begin{pr} Let $i$ be the final iteration. Since the algorithm stopped, we know $K_i=K_{i-1} \leq 3$. The colour is encoded as $<bin(I), bin_I(c)>$ since $I\leq 3$ we get only six possible colours.

For the running time we claim: $K_i\leq 2+ \left\lceil \log^{(i)} K_0 \right\rceil$. This can be proven by induction on $i$. 

$i=1$ \ok. $i\rightarrow i+1$:

\begin{align*}
K_{i+1}&= 1+\rup{\log K_i}\\
	&\leq 1+ \rup{\log\left(2+\rup{\log^{(i)}K_0}\right)}\\
	&\leq 1+\rup{1+\log \rup{\log^{(i)} K_0}}\\
	&\leq 2+\rup{\log \rup{\log^{(i)} K_0}}\\
	&\leq 2+\rup{\log \log^{(i)} K_0}\\
\end{align*}

If we choose $i=\log^*(K_0)$ we get $K_i\leq 4$ by definition of $\log^*$. Hence in round $\log^*(K_0)+2$ the algorithm will finish.
\end{pr}

Six colours isn't very satisfying for a bipartite graph, so we try to reduce the number of colours further.

\begin{figure}[hbt]
\begin{lstlisting}
root: choose a different colour
all others: take the old colour of parent
\end{lstlisting}
\caption{Shift-Down}
\label{alg:Shift-Down}
\end{figure}

Algorithm \ref{alg:Shift-Down} only takes one round.

\begin{lem} The colouring produced by algorithm \ref{alg:Shift-Down} is valid, if the initial colouring is valid. Also, all siblings of a node have the same colour.\end{lem}

\begin{pr} Let $c'_u$ be the initial colours and $c_u$ be the new ones. Let $u_1\rightarrow u_2 \rightarrow u_3$ be a path in the tree. Then $c_{u_2}=c'_{u_1} \neq c_{u_3}=c'_{u_2}$. The root is also ok, since it chooses a fresh colour.\end{pr}

\begin{figure}[hbt]
\begin{lstlisting}
for $x\in \{3,4,5\}$
	Shift-Down
	All vertices coloured with $x$ call First-Free
\end{lstlisting}
\caption{6-to-3}
\label{alg:6-to-3}
\end{figure}

\begin{lem} The colouring produced by \ref{alg:6-to-3} is valid and contains $\leq 3$ colours.\end{lem}

\begin{pr} After the shift down, the neighbourhood of a vertex labelled with $x$ contains only two colours. The colour of the parent and the colours of the children. It can also not happen that two adjacent vertices call First-Free simultaneously, since then they would have had the same colour to begin with.\end{pr}

Finding a colouring with just 2 colours is not possible as fast.